---
title: "Mapping Inequality"
author: "Anna Ramji & Jake McCurdy"
date: "10/28/2020"
output: html_document
---

### From "The ecological and evolutionary consequences of systemic racism" 

**Hello!** In order to understand how to read and interpret these interactive maps, you'll want to understand the metrics that we're using. 

**Holc Grade** systematically grades each neighborhood (as defined by historic trends in **redlining** from 1950) from A to D:

-  **A: "Best"**

-  **B: "Still Desirable"**

-  **C: "Definitely Declining"**

-  **D: "Hazardous"**
 
 Sources:  
- <https://ncrc.org/holc-health/> 
& 
- <https://ncrc.org/holc> 

 
**Normalized Difference Vegetation Index (NDVI)** (raster data)
is used as proxy measure of vegetation health, cover and phenology (life cycle stage) over large areas. It is calculated using multiple bands from satellite images. Values closer to 1 indicate presence of vegetation, while near-zero values indicate water and the range between is a range from high density or presence of vegetation to no vegetation present.  


```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, 
                      warning = FALSE, results = "hide",
                      cache = TRUE)

# install.packages(c("tmap", "raster", "sf")) 
library(tmap)      #interactive maps, raster + vector layers
library(raster)       #Main raster library
library(tidyverse)    #our old friend
library(sf)           #to work with simple features data

tmp <- tempfile()
download.file("https://dsl.richmond.edu/panorama/redlining/static/fullshpfile.zip", tmp)
unzip(tmp)

tmp <- tempfile()
download.file("https://data.ca.gov/dataset/e212e397-1277-4df3-8c22-40721b095f33/resource/b0007416-a325-4777-9295-368ea6b710e6/download/ca-county-boundaries.zip", tmp) 
unzip(tmp)
```

## Birmingham 
### 3 Layers: Holc Grade for each neighborhood, Mean NDVI for each neighborhood, and Total NDVI 
*(shows NDVI for each pixel of area 30 m  x  30 m)*

```{r Birmingham, results="markup", cache = FALSE, echo=FALSE}

holc <- st_read("fullshpfile/shapefile/holc_ad_data.shp")


holc_birmingham <- holc %>%
  filter(city == "Birmingham")

ndvi_birmingham <- raster("../geospatial-mccurdy-ramji/data/NDVI/composite_birmingham.tif")

# Birmingham shp & NDVI
shp_birmingham <- holc %>% filter(city == "Birmingham")

mean_ndvi_birmingham <-
  holc_birmingham %>%
  mutate(mean_ndvi = raster::extract(ndvi_birmingham, holc_birmingham, fun = mean)) 

mean_ndvi_birmingham <- 
  mean_ndvi_birmingham %>% 
  mutate(mean_ndvi = as.numeric(mean_ndvi))

tmap_mode("view")
tm_shape(shp_birmingham) +
  tm_polygons("holc_grade", palette = "Reds", alpha=0.4) +
  tm_shape(ndvi_birmingham) +
  tm_raster(alpha=0.7) +
  tm_shape(mean_ndvi_birmingham) +
  tm_polygons(col = "mean_ndvi")

```

## San Francisco 
### 3 Layers: Holc Grade for each neighborhood, Mean NDVI for each neighborhood, Total NDVI 
*(shows NDVI for each pixel of area 30 m  x  30 m)*  

```{r San Francisco, results="markup", cache = FALSE, echo=FALSE}

holc_SF <- holc %>%
  filter(city == "San Francisco")


shp_SF <- holc %>% filter(city == "San Francisco") %>% select(holc_grade, geometry) 


ndvi_SF <- raster("../geospatial-mccurdy-ramji/data/NDVI/composite_SF.tif")


mean_ndvi_SF <-
  holc_SF %>%
  mutate(mean_ndvi = raster::extract(ndvi_SF, holc_SF, fun = mean)) 


mean_ndvi_SF <- 
  mean_ndvi_SF %>% 
  mutate(mean_ndvi = as.numeric(mean_ndvi))


tmap_mode("view")
tm_shape(shp_SF) +
  tm_polygons("holc_grade", palette = "Reds", alpha=0.4) + 
  tm_shape(ndvi_SF) +
  tm_raster(style = "cont", palette = "Greens", alpha = 0.6) +
  tm_shape(mean_ndvi_SF) +
  tm_polygons(col = "mean_ndvi")

```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
